### 2014-11-09 TransitionsUsefulSecond1.tex
stuff

\inference
[NonMem]
{\theta(i) = (s, \pc) & \dec(s,\pc) = (\nm, x) \\ \exec(s, \pc, (\nm, x)) = (\pc', s')}
{\io{SC}{\theta,m}{\theta[i \coloneqq (s', \pc')], m}{}}

\inference
[Load]
{\theta(i) = (s,\pc) & \dec(s,\pc) = (\ld, x, a) \\ \exec(s, \pc, (\ld, x, m(a))) = (\pc', s')}
{\io{SC}{\theta,m}{\theta[i\coloneqq(s',\pc')],m}{}}

\inference
[Store]
{\theta(i) = (s,\pc) & \dec(s,\pc) = (\st, a, v) \\ \exec(s, \pc, (\st)) = (\pc', s')}
{\io{SC}{\theta,m}{\theta[i\coloneqq(s',\pc')],m[a\coloneqq v]}{}}
\end{boxedminipage}
\caption{LTS for SC with $n$ simple processors}
\label{Ref}
\end{figure}

@@

Figure~\ref{Ref} defines an $n$-processor, sequentially consistent system as an
LTS.  The definition is parametrized over details of an \emph{instruction set
architecture} (ISA).  In particular, the ISA gives us some domains of
architectural states $s$ (e.g., register files) and of program counters $\pc$.
A function $\dec(s, \pc)$ figures out which instruction $\pc$ references in the
current state, returning the instruction's ``decoded'' form.  A companion
function $\exec(s, \pc, \mathit{dec})$ actually executes the instruction,
returning the next program counter $\pc'$ and the new state $s'$.

@@

\inference
[NM]
{\dec(s,\pc) = (\nm, x) & \exec(s, \pc, (\nm, x)) = (\pc', s')}
{\ioe{P$_\text{ref}$}{s,\pc,\bot}{s', \pc', \bot}}

@@

\inference
[LdRp]
{\dec(s,\pc) = (\ld, x, a) & \exec(s, \pc, (\ld, x, v)) = (\pc', s')}
{\iol{P$_\text{ref}$}{s, \pc, \top}
{s', \pc', \bot}{\epsilon, \ld, v}}

\inference
[StRp]
{\dec(s,\pc) = (\st, a, v) & \exec(s, \pc, (\st)) = (\pc', s')}
{\iol{P$_\text{ref}$}{s, \pc, \top}
{s', \pc', \bot}{\st}}
\end{boxedminipage}

@@

\begin{math}
\;\;\;\;\;\left\{
\begin{array}{ll}
s', \pc' &: (\ld, v) \in \mathit{rs} \wedge \dec(pc, s) = (\ld, x, a) \wedge\\
\multicolumn{2}{r}{\exec(pc, s, (\ld, x, v)) = (\pc', s'))}\\
s', \pc' &: (\st) \in \mathit{rs} \wedge \dec(pc, s) = (\st, a, v) \wedge\\
\multicolumn{2}{r}{\exec(pc, s, (\st)) = (\pc', s')}\\
s, \pc &: \text{otherwise if } \dec(pc,s) \neq H\\
H &: \text{otherwise}
\end{array}
\right.
\end{math}
\end{boxedminipage}
\caption{Mapping of states from $(\text{P$^n_\text{ref}$} + M_m)$ to SC}
\label{smap}
\end{figure}

@@

\inference[LdRpBad]
{\commit(\rob) = (\pc,\pc',(\ld,x,a,v',s')) \\ v' \neq v & \exec(s,\pc,(\ld,x,v)) = (\pc'', s'')}
{\iox{\text{P$_\text{so}$}}{s,\pc,\top,\\\rob,\ppc}{s'',\pc'',\bot,\phi,\\\setBp(\ppc,\pc)}{\text{ToP}(\epsilon, \ld, v)}}

@@

\begin{math}
\small
\;\;\;\;\;\;\;\;\left\{
\begin{array}{ll}
\commit(\rob) = (\pc, \pc', (\nm, s')) &\Rightarrow
\exists x. \; \dec(s,\pc) = (\nm, x) \wedge \exec(s,\pc, (\nm,x)) =
(\pc', s')\\
\commit(\rob) = (\pc, \pc', (\ld, x, a, v, s')) & \Rightarrow
\dec(s,\pc) = (\ld, x,a) \wedge \exec(s,\pc, (\ld,x,v)) = (\pc', s')\\
\commit(\rob) = (\pc, \pc', (\st, a, v, s')) &\Rightarrow
\dec(s,\pc) = (\st, a,v) \wedge \exec(s,\pc, (\st)) =
(\pc', s')\\
\commit(\rob) = H & \Rightarrow
\dec(s, \pc) = H\\
\end{array}
\right.\end{math}
\label{rob}
\end{inv}
\vspace{-.5cm}
\caption{Correctness of reorder buffer}
\label{robfig}
\end{figure*}

>>>>

\inference
[NonMem]
{\theta(i) = (s, \pc) & \dec(s,\pc) = (\nm, x) \\ \exec(s, \pc, (\nm, x)) = (s',\pc')}
{\io{SC}{\theta,m}{\theta[i \coloneqq (s', \pc')], m}{}}

\inference
[Load]
{\theta(i) = (s,\pc) & \dec(s,\pc) = (\ld, x, a) \\ \exec(s, \pc, (\ld, x, m(a))) = (s',\pc')}
{\io{SC}{\theta,m}{\theta[i\coloneqq(s',\pc')],m}{}}

\inference
[Store]
{\theta(i) = (s,\pc) & \dec(s,\pc) = (\st, a, v) \\ \exec(s, \pc, (\st)) = (s',\pc')}
{\io{SC}{\theta,m}{\theta[i\coloneqq(s',\pc')],m[a\coloneqq v]}{}}
\end{boxedminipage}
\caption{LTS for SC with $n$ simple processors}
\label{Ref}
\end{figure}

@@

Figure~\ref{Ref} defines an $n$-processor, sequentially consistent system as an
LTS.  The definition is parametrized over details of an \emph{instruction set
architecture} (ISA).  In particular, the ISA gives us some domains of
architectural states $s$ (e.g., register files) and of program counters $\pc$.
A function $\dec(s, \pc)$ figures out which instruction $\pc$ references in the
current state, returning the instruction's ``decoded'' form.  A companion
function $\exec(s, \pc, \mathit{dec})$ actually executes the instruction,
returning the new state $s'$ and the next program counter $\pc'$.

@@

\inference
[NM]
{\dec(s,\pc) = (\nm, x) & \exec(s, \pc, (\nm, x)) = (s', \pc')}
{\ioe{P$_\text{ref}$}{s,\pc,\bot}{s', \pc', \bot}}

@@

\inference
[LdRp]
{\dec(s,\pc) = (\ld, x, a) & \exec(s, \pc, (\ld, x, v)) = (s',\pc')}
{\iol{P$_\text{ref}$}{s, \pc, \top}
{s', \pc', \bot}{\epsilon, \ld, v}}

\inference
[StRp]
{\dec(s,\pc) = (\st, a, v) & \exec(s, \pc, (\st)) = (s',\pc')}
{\iol{P$_\text{ref}$}{s, \pc, \top}
{s', \pc', \bot}{\st}}
\end{boxedminipage}

@@

\begin{math}
\;\;\;\;\;\left\{
\begin{array}{ll}
s', \pc' &: (\ld, v) \in \mathit{rs} \wedge \dec(pc, s) = (\ld, x, a) \wedge\\
\multicolumn{2}{r}{\exec(pc, s, (\ld, x, v)) = (s',\pc'))}\\
s', \pc' &: (\st) \in \mathit{rs} \wedge \dec(pc, s) = (\st, a, v) \wedge\\
\multicolumn{2}{r}{\exec(pc, s, (\st)) = (s',\pc')}\\
s, \pc &: \text{otherwise if } \dec(pc,s) \neq H\\
H &: \text{otherwise}
\end{array}
\right.
\end{math}
\end{boxedminipage}
\caption{Mapping of states from $(\text{P$^n_\text{ref}$} + M_m)$ to SC}
\label{smap}
\end{figure}

@@

\inference[LdRpBad]
{\commit(\rob) = (\pc,\pc',(\ld,x,a,v',s')) \\ v' \neq v & \exec(s,\pc,(\ld,x,v)) = (s'',\pc'')}
{\iox{\text{P$_\text{so}$}}{s,\pc,\top,\\\rob,\ppc}{s'',\pc'',\bot,\phi,\\\setBp(\ppc,\pc)}{\text{ToP}(\epsilon, \ld, v)}}

@@

\begin{math}
\small
\;\;\;\;\;\;\;\;\left\{
\begin{array}{ll}
\commit(\rob) = (\pc, \pc', (\nm, s')) &\Rightarrow
\exists x. \; \dec(s,\pc) = (\nm, x) \wedge \exec(s,\pc, (\nm,x)) =
(s',\pc')\\
\commit(\rob) = (\pc, \pc', (\ld, x, a, v, s')) & \Rightarrow
\dec(s,\pc) = (\ld, x,a) \wedge \exec(s,\pc, (\ld,x,v)) = (s',\pc')\\
\commit(\rob) = (\pc, \pc', (\st, a, v, s')) &\Rightarrow
\dec(s,\pc) = (\st, a,v) \wedge \exec(s,\pc, (\st)) =
(s',\pc')\\
\commit(\rob) = H & \Rightarrow
\dec(s, \pc) = H\\
\end{array}
\right.\end{math}
\label{rob}
\end{inv}
\vspace{-.5cm}
\caption{Correctness of reorder buffer}
\label{robfig}
\end{figure*}

### 2014-11-09 TransitionsUsefulThird1.tex
stuff

We believe that other invalidation-based protocols with inclusive caches (\ie{}
those cache hierarchies where a parent cache contains all the locations that
any of its children contains) like MESI, MOSI, and MOESI also satisfy our
lemmas, making their verification easier by reusing most of our MSI proof; we
omit the discussion of extension of this proof to other protocols for space
reasons.

@@

\begin{lemma}
\textit{nonAncestorCompatible}: For two distinct caches $c_1$ and $c_2$, neither of which is the
ancestor of the other in the cache-tree,
%such
%that neither is an ancestor of the other in the cache-tree hierarchy (where
%the ancestor relation is the reflexive-transitive closure of the parent relation),
coherence states $\s(c_1,a)$ and $\s(c_2,a)$ are
\textit{compatible} (defined below) for every address $a$.
\label{nonAncestorCompatible}
\end{lemma}
\begin{defn}
\textit{compatible}: Coherence states $x_1$ and $x_2$ are \textit{compatible}
iff whenever either $x_1$ or $x_2$ is \Mo{}, the other is \In.
\label{compatible}
\end{defn}

\begin{lemma}
\textit{noTransitWrite}: Whenever data for an address $a$ is in transit (\ie{}
$\forall T. \; T_s \le T \le T_r$ where $T_s$ is the time of sending the data and
$T_r$ the time of receiving the data), no cache can process a store request for
$a$, and the data must be sent from a \textit{clean} cache.
\label{noTransitWrite}
\end{lemma}

\begin{lemma}
\textit{conservative}: A cache's state for an address is never greater than
the corresponding state in its parent's directory.
\label{conservative}
\end{lemma}

\begin{lemma}
\textit{localCompatibility1}:
The state of a cache is never less than its directory entry for any of
its children.
\label{localCompatibility1}
\end{lemma}
\begin{lemma}
\textit{localCompatibility2}: Let $c_1$ and $c_2$ be distinct caches with the
same parent $p$. Then, $\s(c_1, a)$ and $\s(c_2, a)$ are
\textit{compatible}.
\label{localCompatibility2}
\end{lemma}

>>>>

@@

%\begin{lemma}
%\textit{nonAncestorCompatible}: For two distinct caches $c_1$ and $c_2$, neither of which is the
%ancestor of the other in the cache-tree,
%%such
%%that neither is an ancestor of the other in the cache-tree hierarchy (where
%%the ancestor relation is the reflexive-transitive closure of the parent relation),
%coherence states $\s(c_1,a)$ and $\s(c_2,a)$ are
%\textit{compatible} (defined below) for every address $a$.
%\label{nonAncestorCompatible}
%\end{lemma}
%\begin{defn}
%\textit{compatible}: Coherence states $x_1$ and $x_2$ are \textit{compatible}
%iff whenever either $x_1$ or $x_2$ is \Mo{}, the other is \In.
%\label{compatible}
%\end{defn}

\begin{lemma}[noTransitWrite] Whenever data for an address $a$ is in transit (\ie{}
$\forall T. \; T_s \le T \le T_r$ where $T_s$ is the time of sending the data and
$T_r$ the time of receiving the data), no cache can process a store request for
$a$, and the data must be sent from a \textit{clean} cache.
\label{noTransitWrite}
\end{lemma}

\begin{lemma}%[Conservative directory] 
$\forall p, \forall c, \forall a\;\s(c, a) \le \dst(p, c, a)$
\label{conservative}
\end{lemma}

\begin{lemma}%[Compatible directory]
$\forall p, \forall c, \forall a.\; \compat(p, c, \dst(p,c,a), a)$
\end{lemma}

\begin{lemma}%[Coherence state vs Directory state]
$\forall p, \forall c, \forall a.\; \dst(p,c,a) \le \s(p, a)$
\end{lemma}
%\begin{lemma}
%\textit{localCompatibility1}:
%The state of a cache is never less than its directory entry for any of
%its children.
%\label{localCompatibility1}
%\end{lemma}
%\begin{lemma}
%\textit{localCompatibility2}: Let $c_1$ and $c_2$ be distinct caches with the
%same parent $p$. Then, $\s(c_1, a)$ and $\s(c_2, a)$ are
%\textit{compatible}.
%\label{localCompatibility2}
%\end{lemma}

The same proof structure can be used to prove other invalidation-based
protocols with inclusive caches (\ie{} those cache hierarchies in which a
parent cache contains all the locations that any of its children contains) like
MESI, MOSI, and MOESI. We omit the discussion of extension of this proof to
other protocols for space reasons.
